Design Document: Monte Carlo Maximum Likelihood for Generalized Linear Mixed Models

### 2015-01-09 NewDesignDoc/DesignDocGLMM.tex
Continuing proof. About half done.

\section{Central Limit Theorem for MCLA}\label{sec:CLT}
Define
\begin{align}
\gamma_1&= \int f_\theta(u,y) du\\
\gamma_2 &=\int \tilde{f}(u)du.
\end{align}
Recall the calculation for the MCLA gradient originally stated in  \eqref{eq:MCLAgradient}:
\begin{align}
\nabla l_m(\theta)&= \dfrac{\sum_{k=1}^m    \left( \nabla \log f_\theta(u_k,y) - \nabla \log \tilde{f} (u_k)  \right) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\sum_{k=1}^m \left( \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right) }\\
&=\dfrac{\sum_{k=1}^m     \nabla \log f_\theta(u_k,y)   \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\sum_{k=1}^m \left( \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right) }   -\dfrac{\sum_{k=1}^m     \nabla \log \tilde{f}(u_k)   \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\sum_{k=1}^m \left( \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right) }
\end{align}

@@

\dfrac{1}{m}\sum_{k=1}^m \nabla \log \tilde{f}(u_k)  \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} &\rightarrow  E_{\tilde{f}} \left[ \nabla \log \tilde{f}(u)  \dfrac{f_\theta(u,y)}{\tilde{f}(u)}  \right] \\
&= \int \nabla \log \tilde{f}(u)  \dfrac{f_\theta(u,y)}{\tilde{f}(u)} \dfrac{\tilde{f}(u)}{\gamma_2} du\\
&= \dfrac{\gamma_1}{\gamma_2} \int \nabla \log \tilde{f}(u)  \dfrac{f_\theta(u,y)}{\gamma_1}  du \\
&=\dfrac{\gamma_1}{\gamma_2} E_f \left[ \nabla \log \tilde{f}(u)  \right]

@@

Then, by Slutsky's,

@@

\nabla l_m(\theta)&\rightarrow \dfrac{  \dfrac{\gamma_1}{\gamma_2} E_f \left[ \nabla \log f_\theta(u,y)  \right] }{\dfrac{\gamma_1}{\gamma_2}}-\dfrac{  \dfrac{\gamma_1}{\gamma_2} E_f \left[ \nabla \log \tilde{f}(u)  \right] }{\dfrac{\gamma_1}{\gamma_2}}\\
&=  E_f \left[ \nabla \log f_\theta(u,y) - \nabla \log \tilde{f}(u)  \right] \\

@@

&\sqrt{m} \left[  \dfrac{\dfrac{1}{m}\sum_{k=1}^m    \left( \nabla \log f_\theta(u_k,y) - \nabla \log \tilde{f}(u_k) \right)  \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\dfrac{1}{m} \sum_{k=1}^m \left( \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right) }- l(\theta)  \right] \\

@@

\dfrac{\dfrac{1}{\sqrt{m}}\sum_{k=1}^m    \left( \nabla \log f_\theta(u_k,y)  - \nabla \log \tilde{f}(u_k) \right) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\dfrac{1}{m} \sum_{k=1}^m \left( \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right) }- \sqrt{m}\;  l(\theta)  \\
&=\dfrac{\dfrac{1}{\sqrt{m}}\sum_{k=1}^m    \left( \nabla \log f_\theta(u_k,y) - \nabla \log \tilde{f}(u_k) \right) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} 

@@

\sum_{k=1}^m     \nabla \log \tilde{f}(u_k)\dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} 
are each finite.

@@

\sum_{k=1}^m \dfrac{f_\theta(u_k)}{\tilde{f}(u_k)}

@@

\int \dfrac{f_\theta(u)^2}{\tilde{f}(u)} du < \infty.

@@

\int \dfrac{f_\theta(u)^2}{\tilde{f}(u)} du &= \int \dfrac{N(u|0,D)^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du \\
& \leq \int \dfrac{N(u|0,D)^2}{p_1 N(u|0,D)} du \\
&= \dfrac{1}{p_1} \int N(u|0,D) du \\
&= \dfrac{1}{p_1}.

@@

Therefore $\sum_{k=1}^m f_\theta(u_k)$ has finite variance as long as $p_1\neq 0$.

Second, we want to show that
\begin{align}
\sum_{k=1}^m \nabla \log f_\theta (u_k,y) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}
\end{align}
 has finite variance. That is, we want to show the existence of
\begin{align}
\int \dfrac{N(u|0,D)^2 \left[\nabla \left(\log f_\theta(y|u)+\log f_\theta (u) \right) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du.
\end{align}
That is, we want to show the existence of
\begin{align}
\int &\dfrac{N(u|0,D)^2 \left[\nabla \left(\log f_\theta(y|u)+\log f_\theta (u) \right) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du \\
&\leq \int  \dfrac{N(u|0,D)^2 \left[\nabla \left(\log f_\theta(y|u) + \log f_\theta (u) \right) \right]^2}{p_1 N(u|0,D)} du  \\
&\leq  \int \dfrac{N(u|0,D)^2 \left[ \dfrac{\partial}{\partial \beta} \log f_\theta(y|u) \, \, \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2}{p_1 N(u|0,D)} du  \\
&\leq \dfrac{1}{p_1}  \int N(u|0,D) \left[ \dfrac{\partial}{\partial \beta}\log f_\theta(y|u) \, \, \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2 du.
\end{align}
By the Cauchy-Schwartz inequality, 
\begin{align}
  &\int N(u|0,D) \left[ \dfrac{\partial}{\partial \beta}\log f_\theta(y|u) \, \, \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2 du \\
& \leq \left[  \int N(u|0,D) \left[  \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2 du \right]^{1/2}  \left[  \int N(u|0,D) \left[ \dfrac{\partial}{\partial \beta}\log f_\theta(y|u)   \right]^2 du \right]^{1/2} .
\end{align}
Therefore, I need to show that each of these two integrals is finite.  I will start with the first of the two integrals: 
\begin{align}
\int N(u|0,D) \left[  \dfrac{\partial}{\partial \nu} \log f_\theta(u)  \right]^2 du. 
\end{align}
Recall the notation from section \ref{sec:Ddiag}.  If
\begin{align}
\int N(u_t|0,D_t) \left[  \dfrac{\partial}{\partial \nu_t} \log f_\theta(u)  \right]^2 du_t< \infty 
\end{align}
for every $t=1,...,T$, then
\begin{align}
\int N(u|0,D) \left[  \dfrac{\partial}{\partial \nu} \log f_\theta(u)  \right]^2 du < \infty
\end{align}
by Cauchy Schwartz.  Therefore, I will prove the integral is finite for the general case of $u_t$ and $\nu_t$. First, we recognize that our integral is actually an expectation of a function of a normal random variable:
\begin{align}
\int N(u_t|0,D_t) \left[  \dfrac{\partial}{\partial \nu_t} \log f_\theta(u)  \right]^2 du_t &=
\int \dfrac{1}{\nu_t^{q_t/2}} e^{-u_t'u_t/(2 \nu_t)} \left[ -\dfrac{q_t}{2 \nu_t} + \dfrac{u_t'u_t}{2 \nu_t^2}  \right]^2 du_t .
\end{align}
Normal distributions have finite moments of all orders, so we know that\begin{align}
\int N(u_t|0,D_t) \left[  \dfrac{\partial}{\partial \nu_t} \log f_\theta(u)  \right]^2 du_t < \infty.
\end{align}

@@

Third and lastly, we want to show that 
\begin{align}
\sum_{k=1}^m \nabla \log \tilde{f} (u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}
\end{align}
 has finite variance. That is, we want to show the existence of
\begin{align}
\int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du.
\end{align}
We see that
\begin{align}
\int& \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du\\
&\leq  \int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)} du\\
&= \dfrac{1}{p_1} \int N(u|0,D) \left[\nabla \log \tilde{f}u) \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)+p_2 N(u|u^*,D^*)+p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})}   \right]^2 du\\
&\leq\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)}   \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{ \dfrac{\partial}{\partial \nu} N(u|0,D)}{ N(u|0,D)}   \right]^2 du\\
&= \dfrac{1}{p_1} \int \dfrac{\left[ \dfrac{\partial}{\partial \nu} N(u|0,D)  \right]^2}{N(u|0,D)} du \\
\end{align}
Now we use
\begin{align}
\dfrac{\partial}{\partial \nu} \log N(u|0,D) = \dfrac{\dfrac{\partial}{\partial \nu}  N(u|0,D)}{N(u|0,D)}
\end{align}
to see
\begin{align}
 \dfrac{1}{p_1} \int \dfrac{\left[ \dfrac{\partial}{\partial \nu} N(u|0,D)  \right]^2}{N(u|0,D)} du &=\dfrac{1}{p_1} \int  N(u|0,D) \left[ \dfrac{\partial}{\partial \nu} \log N(u|0,D)  \right]^2 du \\
&= \dfrac{1}{p_1} \int  N(u|0,D) \left[ \dfrac{\partial}{\partial \nu} \log N(u|0,D)  \right]^2 du 
\end{align}
which exists and is finite.

>>>>

\section{Central Limit Theorem for MCLA}\label{sec:CLT}
In this section, we focus on the gradient of the MCLA because the trust criterion for finding the maximum is based on the gradient.
Define
\begin{align}
\gamma_1&= \int f_\theta(u,y) du\\
\gamma_2 &=\int \tilde{f}(u)du.
\end{align}
Recall the calculation for the MCLA gradient originally stated in  \eqref{eq:MCLAgradient}:
\begin{align}
\nabla l_m(\theta)&= \dfrac{\sum_{k=1}^m     \nabla \log f_\theta(u_k,y)  \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\sum_{k=1}^m \left( \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right) }
\end{align}

@@

%and
%\begin{align}
%\dfrac{1}{m}\sum_{k=1}^m \nabla \log \tilde{f}(u_k)  \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} &\rightarrow  E_{\tilde{f}} \left[ \nabla \log \tilde{f}(u)  \dfrac{f_\theta(u,y)}{\tilde{f}(u)}  \right] \\
%&= \int \nabla \log \tilde{f}(u)  \dfrac{f_\theta(u,y)}{\tilde{f}(u)} \dfrac{\tilde{f}(u)}{\gamma_2} du\\
%&= \dfrac{\gamma_1}{\gamma_2} \int \nabla \log \tilde{f}(u)  \dfrac{f_\theta(u,y)}{\gamma_1}  du \\
%&=\dfrac{\gamma_1}{\gamma_2} E_f \left[ \nabla \log \tilde{f}(u)  \right]
%\end{align}

@@

Then, by Slutsky's theorem,

@@

\nabla l_m(\theta)&\rightarrow \dfrac{  \dfrac{\gamma_1}{\gamma_2} E_f \left[ \nabla \log f_\theta(u,y)  \right] }{\dfrac{\gamma_1}{\gamma_2}}\\
&=  E_f \left[ \nabla \log f_\theta(u,y) \right] \\

@@

&\sqrt{m} \left[  \dfrac{\dfrac{1}{m}\sum_{k=1}^m    \nabla \log f_\theta(u_k,y)   \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\dfrac{1}{m} \sum_{k=1}^m  \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}  }- l(\theta)  \right] \\

@@

\dfrac{\dfrac{1}{\sqrt{m}}\sum_{k=1}^m     \nabla \log f_\theta(u_k,y)   \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }{\dfrac{1}{m} \sum_{k=1}^m  \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} }- \sqrt{m}\;  l(\theta)  \\
&=\dfrac{\dfrac{1}{\sqrt{m}}\sum_{k=1}^m     \nabla \log f_\theta(u_k,y)   \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} 

@@

are finite.

@@

@@

\int \dfrac{f_\theta(u,y)^2}{\tilde{f}(u)} du < \infty.

@@

\int \dfrac{f_\theta(u,y)^2}{\tilde{f}(u)} du &=\int \dfrac{f_\theta(y|u)^2 f_\theta(u)^2}{\tilde{f}(u)} du\\
&\leq \int \dfrac{f_\theta(y|u)^2 f_\theta(u)^2}{p_1 \grave{f}(u)} du\\
&\propto \dfrac{1}{p_1} \int \dfrac{\exp(-u'D^{-1}u) \, f_\theta(y|u)^2  }{\left[ 1+u'u/\zeta  \right]^{-(\zeta+q)/2}} du

@@

Since
f_\theta(y|u)^2=\left( e^{y'\eta-c(\eta)} \right)^2 = e^{2y'\eta-2c(\eta)} ,
\int \dfrac{f_\theta(u,y)^2}{\tilde{f}(u)} du &\propto \dfrac{1}{p_1} \int \dfrac{\exp(-u'D^{-1}u) \, \exp(2y'\eta)  }{\left[ 1+u'u/\zeta  \right]^{-(\zeta+q)/2} \,  \exp(2c(\eta))  } du\\
&=\dfrac{1}{p_1} \int \dfrac{ \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} \, \exp(2y'\eta)  }{\exp(u'D^{-1}u)   \,  \exp(2c(\eta))  } du \\
&=\dfrac{1}{p_1} \int \dfrac{ \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} \, \exp(2y'Zu)  }{\exp(u'D^{-1}u)   \,  \exp(2c(\eta))  } du 
When $y|u\sim$ Bernoulli, then
c(\eta)=\log(1+e^\eta) \Rightarrow \exp(2c(\eta))&= \exp( 2 \log(1+e^\eta)  )\\
&= ( \exp( \log(1+e^\eta)  )  )^2\\
&= (1+e^\eta) ^2
\int \dfrac{f_\theta(u,y)^2}{\tilde{f}(u)} du &\propto \dfrac{1}{p_1} \int \dfrac{ \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} \, \exp(2y'Zu)  }{\exp(u'D^{-1}u)   \,  (1+e^\eta) ^2  } du,
which converges since $\exp(u'D^{-1}u$ grows faster than any term in the numerator of the integrand.

On the other hand, when $y|u\sim$ Poisson, then 
\begin{align}
c(\eta)=\exp(\eta) \Rightarrow \exp(2 c(\eta)) = (\exp(\exp(\eta)))^2
\end{align}
and
\begin{align}
\int \dfrac{f_\theta(u,y)^2}{\tilde{f}(u)} du &\propto \dfrac{1}{p_1} \int \dfrac{ \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} \, \exp(2y'Zu)  }{\exp(u'D^{-1}u)   \,   (\exp(\exp(\eta)))^2  } du,
\end{align}
which converges since  $(\exp(\exp(\eta)))^2$ grows faster than any term in the numerator of the integrand.

Therefore, whether the data are Bernoulli or Poisson distributed, the importance sampling weights 
\begin{align}
\sum_{k=1}^m  \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}
\end{align}
have finite variance.

Now, we want to show that
\begin{align}
\sum_{k=1}^m \nabla \log f_\theta (u_k,y) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} =\sum_{k=1}^m \nabla \log f_\theta (y|u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} +\sum_{k=1}^m \nabla \log f_\theta (u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} 
\end{align}
 has finite variance. We need to show the following are finite:
\begin{align}
\var \left( \sum_{k=1}^m \nabla \log f_\theta (y|u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}   \right) \\
\var \left(\sum_{k=1}^m \nabla \log f_\theta (u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right)\\
\cov \left(\sum_{k=1}^m \nabla \log f_\theta (y|u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \, , \, \sum_{k=1}^m \nabla \log f_\theta (u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)} \right)_.
\end{align}

That is, we want to show the existence of the following:
\begin{align}
&\int \dfrac{ \left[ \nabla \log f_\theta (y|u) \right]^2 \left[ f_\theta(u,y)\right]^2}{\tilde{f}(u)} du \label{eq:first} \\
&\int \dfrac{ \left[ \nabla \log f_\theta (u) \right]^2 \left[ f_\theta(u,y)\right]^2}{\tilde{f}(u)} du \label{eq:second} \\
&\int \dfrac{  \nabla \log f_\theta (u)  \nabla \log f_\theta (y|u)  \left[ f_\theta(u,y)\right]^2}{\tilde{f}(u)} du \\
\end{align}

First, we start with \eqref{eq:first}:
\begin{align}
\int \dfrac{ \left[ \nabla \log f_\theta (y|u) \right]^2 \left[ f_\theta(u,y)\right]^2}{\tilde{f}(u)} du 
&=\int \dfrac{ \left[ \nabla \log f_\theta (y|u) \right]^2 \left[ f_\theta(y|u)\right]^2 \left[ f_\theta(u)\right]^2}{\tilde{f}(u)} du \\
&\leq \int \dfrac{ \left[ \nabla \log f_\theta (y|u) \right]^2 \left[ f_\theta(y|u)\right]^2 \left[ f_\theta(u)\right]^2}{p_1\grave{f}(u)} du \\
&\propto \int \dfrac{ \left[ \nabla \log f_\theta (y|u) \right]^2 \left[ f_\theta(y|u)\right]^2 \exp(-u'D^{-1}u)}{p_1\left[ 1+u'u/\zeta  \right]^{-(\zeta+q)/2}} du \\
&= \int \dfrac{ \left[ \nabla \log f_\theta (y|u) \right]^2 \left[ f_\theta(y|u)\right]^2 \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} }{p_1 \exp(u'D^{-1}u)} du \\
&= \int \dfrac{ \left[ \nabla \log f_\theta (y|u) \right]^2 \, \exp(2y'(X \beta+Zu)) \,  \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} }{ p_1 \, \exp(2 c(\eta)) \,\exp(u'D^{-1}u)} du \label{eq:soon}
\end{align}
By the Cauchy-Schwartz inequality, the integral in \eqref{eq:soon} exists if
\begin{align}
 & \int \dfrac{ \left[\dfrac{\partial}{\partial \beta} \log f_\theta (y|u) \right]^2 \, \exp(2y'(X \beta+Zu)) \,  \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} }{ p_1 \, \exp(2 c(\eta)) \,\exp(u'D^{-1}u)} du \nonumber \\
&\propto  \int \dfrac{ \left[X' c'(\eta) \right]^2 \, \exp(2y'(X \beta+Zu)) \,  \left[ 1+u'u/\zeta  \right]^{(\zeta+q)/2} }{ p_1 \, \exp(2 c(\eta)) \,\exp(u'D^{-1}u)} du \label{eq:soon2}
\end{align}
exists, where $c'(\eta)$ is a vector with components $c'(\eta_i)$.  This will converge as long as no term in the numerator of the integrand grows more quickly than $\exp(u'D^{-1}u$. The second and third terms grow less quickly than $\exp(u'D^{-1}u$, so $\left[ c'(\eta)\right]^2$ is the only term to check.  If $y|u \sim$ Bernoulli, then 
\begin{align}
c'(\eta_i)= \dfrac{\exp(\eta_i)}{1+\exp(\eta_i)},
\end{align}
which does not grow more quickly than $\exp(u'D^{-1}u)$. If $y|u \sim$ Poisson, then
\begin{align}
c'(\eta_i)= e^{\eta_i},
\end{align}
which does not grow more quickly than $\exp(u'D^{-1}u)$. Therefore, \eqref{eq:first} exists and is finite, and the variance of 
\begin{align}
\sum_{k=1}^m \nabla \log f_\theta (y|u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}
\end{align}
is finite.

Next, we move on to showing the existence of \eqref{eq:second}.
EDITED TO HERE.

 That is, we want to show the existence of
\begin{align}
\int \dfrac{N(u|0,D)^2 \left[\nabla \left(\log f_\theta(y|u)+\log f_\theta (u) \right) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du.
\end{align}
That is, we want to show the existence of
\begin{align}
\int &\dfrac{N(u|0,D)^2 \left[\nabla \left(\log f_\theta(y|u)+\log f_\theta (u) \right) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du \\
&\leq \int  \dfrac{N(u|0,D)^2 \left[\nabla \left(\log f_\theta(y|u) + \log f_\theta (u) \right) \right]^2}{p_1 N(u|0,D)} du  \\
&\leq  \int \dfrac{N(u|0,D)^2 \left[ \dfrac{\partial}{\partial \beta} \log f_\theta(y|u) \, \, \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2}{p_1 N(u|0,D)} du  \\
&\leq \dfrac{1}{p_1}  \int N(u|0,D) \left[ \dfrac{\partial}{\partial \beta}\log f_\theta(y|u) \, \, \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2 du.
\end{align}
By the Cauchy-Schwartz inequality, 
\begin{align}
  &\int N(u|0,D) \left[ \dfrac{\partial}{\partial \beta}\log f_\theta(y|u) \, \, \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2 du \\
& \leq \left[  \int N(u|0,D) \left[  \dfrac{\partial}{\partial \nu} \log f_\theta (u)  \right]^2 du \right]^{1/2}  \left[  \int N(u|0,D) \left[ \dfrac{\partial}{\partial \beta}\log f_\theta(y|u)   \right]^2 du \right]^{1/2} .
\end{align}
Therefore, I need to show that each of these two integrals is finite.  I will start with the first of the two integrals: 
\begin{align}
\int N(u|0,D) \left[  \dfrac{\partial}{\partial \nu} \log f_\theta(u)  \right]^2 du. 
\end{align}
Recall the notation from section \ref{sec:Ddiag}.  If
\begin{align}
\int N(u_t|0,D_t) \left[  \dfrac{\partial}{\partial \nu_t} \log f_\theta(u)  \right]^2 du_t< \infty 
\end{align}
for every $t=1,...,T$, then
\begin{align}
\int N(u|0,D) \left[  \dfrac{\partial}{\partial \nu} \log f_\theta(u)  \right]^2 du < \infty
\end{align}
by Cauchy Schwartz.  Therefore, I will prove the integral is finite for the general case of $u_t$ and $\nu_t$. First, we recognize that our integral is actually an expectation of a function of a normal random variable:
\begin{align}
\int N(u_t|0,D_t) \left[  \dfrac{\partial}{\partial \nu_t} \log f_\theta(u)  \right]^2 du_t &=
\int \dfrac{1}{\nu_t^{q_t/2}} e^{-u_t'u_t/(2 \nu_t)} \left[ -\dfrac{q_t}{2 \nu_t} + \dfrac{u_t'u_t}{2 \nu_t^2}  \right]^2 du_t .
\end{align}
Normal distributions have finite moments of all orders, so we know that\begin{align}
\int N(u_t|0,D_t) \left[  \dfrac{\partial}{\partial \nu_t} \log f_\theta(u)  \right]^2 du_t < \infty.
\end{align}

@@

Next, we want to show that 
\begin{align}
\sum_{k=1}^m \nabla \log \tilde{f} (u_k) \dfrac{f_\theta(u_k,y)}{\tilde{f}(u_k)}
\end{align}
 has finite variance. That is, we want to show the existence of
\begin{align}
\int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du.
\end{align}
We see that
\begin{align}
\int& \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)+p_2 N(u|u^*, D^*) + p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})} du\\
&\leq  \int \dfrac{N(u|0,D)^2 \left[\nabla \log \tilde{f}u) \right]^2}{p_1 N(u|0,D)} du\\
&= \dfrac{1}{p_1} \int N(u|0,D) \left[\nabla \log \tilde{f}u) \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)+p_2 N(u|u^*,D^*)+p_3 N(u|u^*, ((Z'  c''(X \beta^*+Zu^*) Z +(D^*)^{-1}   )^{-1})}   \right]^2 du\\
&\leq\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{p_1 \dfrac{\partial}{\partial \nu} N(u|0,D)}{p_1 N(u|0,D)}   \right]^2 du\\
&=\dfrac{1}{p_1} \int N(u|0,D) \left[  \dfrac{ \dfrac{\partial}{\partial \nu} N(u|0,D)}{ N(u|0,D)}   \right]^2 du\\
&= \dfrac{1}{p_1} \int \dfrac{\left[ \dfrac{\partial}{\partial \nu} N(u|0,D)  \right]^2}{N(u|0,D)} du \\
\end{align}
Now we use
\begin{align}
\dfrac{\partial}{\partial \nu} \log N(u|0,D) = \dfrac{\dfrac{\partial}{\partial \nu}  N(u|0,D)}{N(u|0,D)}
\end{align}
to see
\begin{align}
 \dfrac{1}{p_1} \int \dfrac{\left[ \dfrac{\partial}{\partial \nu} N(u|0,D)  \right]^2}{N(u|0,D)} du &=\dfrac{1}{p_1} \int  N(u|0,D) \left[ \dfrac{\partial}{\partial \nu} \log N(u|0,D)  \right]^2 du \\
&= \dfrac{1}{p_1} \int  N(u|0,D) \left[ \dfrac{\partial}{\partial \nu} \log N(u|0,D)  \right]^2 du 
\end{align}
which exists and is finite.
